---
title: "CourseProject1"
author: "Kelly"
date: "January 19, 2018"
output: 
    html_document: 
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## What is the mean total number of steps taken per day?

```{r firstchunk}
ActivityData <- read.csv(file="activity.csv",head=TRUE)
library(dplyr)
library(ggplot2)

ActivityByDay <- ActivityData %>% group_by(date) %>% summarize(TotalSteps=sum(steps,na.rm=TRUE))

hist(ActivityByDay$TotalSteps)

MeanByDay <- mean(ActivityByDay$TotalSteps)
MedianByDay <- median(ActivityByDay$TotalSteps)

```

The mean total steps taken per day is `r MeanByDay` and the median total steps taken per day is `r MedianByDay`.

## What is the average daily activity pattern?
```{r secondchunk}
ActivityByInterval <- ActivityData %>% group_by(interval) %>% summarize(MeanSteps=mean(steps,na.rm=TRUE))
plot(ActivityByInterval$interval, ActivityByInterval$MeanSteps, type ="l")

MaxStepsInterval <- ActivityByInterval %>% filter(MeanSteps == max(MeanSteps))


```

The 5-minute interval with the most steps on average is `r MaxStepsInterval$interval`.

## Imputing missing values

```{r thirdchunk}
MissingValues <- is.na(ActivityData)
TotalMissing <- sum(MissingValues)
ActivityDataCopy <- ActivityData

MissingValuesOneCol <- MissingValues[,1]
ActivityDataNAs <- ActivityDataCopy[MissingValuesOneCol,][,1:3]
ActivityDataGood <- ActivityDataCopy[!MissingValuesOneCol,][,1:3]

MergedDataNAs <- merge(ActivityDataNAs, ActivityByInterval, by = c("interval"), all.X = TRUE)
ReplacedData <- MergedDataNAs[,c(4,3,1)]

ColumnNames <- colnames(ActivityDataGood)
colnames(ReplacedData) <- ColumnNames
AllActivityDataReplacements <- rbind(ReplacedData, ActivityDataGood)

ActivityByDayNew <- AllActivityDataReplacements %>% group_by(date) %>% summarize(TotalStepsNew=sum(steps,na.rm=TRUE))

hist(ActivityByDayNew$TotalStepsNew)

MeanByDayNew <- mean(ActivityByDayNew$TotalStepsNew)
MedianByDayNew <- median(ActivityByDayNew$TotalStepsNew)

```

The total number of missing values in the dataset is `r TotalMissing`.

After replacing the missing values with the averages for that time interval from the other days, the mean total steps taken per day is `r MeanByDayNew` and the median total steps taken per day is `r MedianByDayNew`. Yes these values differ from the estimates before the missing values were replaced, and the new totals per day are higher. 


## Are there differences in activity patterns between weekdays and weekends?

```{r fourth chunk}

AllActivityDataReplacements$date <- as.Date(AllActivityDataReplacements$date, format="%Y-%m-%d")
AllActivityDataReplacements$dayofweek <- weekdays(AllActivityDataReplacements$date)

AllActivityDataReplacements$DayType <- ifelse(AllActivityDataReplacements$dayofweek == "Monday" | AllActivityDataReplacements$dayofweek == "Tuesday" | AllActivityDataReplacements$dayofweek == "Wednesday" | AllActivityDataReplacements$dayofweek == "Thursday" | AllActivityDataReplacements$dayofweek == "Friday","WeekDay","WeekEnd")
AllActivityDataReplacements$DayType <- as.factor(AllActivityDataReplacements$DayType)

AllActivityByIntervalDayType <- AllActivityDataReplacements %>% group_by(interval,DayType) %>% summarize(MeanSteps=mean(steps, na.rm=TRUE))

ggplot(AllActivityByIntervalDayType, aes(interval,MeanSteps))+geom_line()+facet_wrap(~DayType,nrow=2)
```
